Speed and strength of an epidemic intervention

### 2020-02-02 doc/speedstrength.tex
tryin to rewrite discussion

The effectiveness of an epidemic intervention is often measured by its ability to reduce the reproductive number below 1.
In this study, we extend this idea by first defining the strength of an intervention on the same scale as the strength of an epidemic (i.e., basic reproductive number $\mathcal R_0$); and then constructing a parallel interpretation which measures the speed of an intervention on the same scale as the speed of an epidemic (i.e., exponential growth rate $r$).
We show that an intervention can successfully contain an epidemic if either its strength or speed is greater than their corresponding measures of the epidemic.

We compare two HIV intervention strategies (condom intervention and test-and-treat intervention) to demonstrate that both strength- and speed-based frameworks can be useful in evaluating the effectiveness of an epidemic intervention.
In particular, we recapitulate the result of \cite{eaton2014proportion} who used detailed mathematical modeling of HIV transmission to show that the amount of early transmission does not affect the effectiveness of the ART.
While they gave a strength-based argument for the result --  increasing the amount of early transmission decreases the basic reproductive number, which negatively correlates with the outcome of the ART intervention \citep{eaton2014proportion} -- there is a simple speed-based intuition behind this phenomenon: in principle, we can control an outbreak if we can identify infected individuals and treat them with ART faster than the \emph{observed} rate at which new cases are generated, which does not depend on the amount of early transmission.
While both strength- and speed-based frameworks give the same conclusion about the outcome of an intervention, one can provide a clearer interpretation of the outcome than the other, depending on the situation.

>>>>

The effectiveness of an epidemic intervention is often measured by its ability to reduce the reproductive number below 1.
In this study, we extended this idea by first defining the strength of an intervention on the same scale as the strength of an epidemic (i.e., basic reproductive number $\mathcal R_0$); and then constructing a parallel interpretation which measures the speed of an intervention on the same scale as the speed of an epidemic (i.e., exponential growth rate $r$).
We showed that an intervention can successfully contain an epidemic if either its strength or speed is greater than their corresponding measures of the epidemic.

We compared two HIV intervention strategies (condom intervention and test-and-treat intervention) to demonstrate that both strength- and speed-based frameworks can be useful in evaluating the effectiveness of an epidemic intervention.
In particular, we provided an alternative explanation for the result of \cite{eaton2014proportion} who used detailed mathematical modeling of HIV transmission to show that the amount of early transmission does not affect the effectiveness of the ART:
we can control an outbreak if we can identify infected individuals and enroll them on ART faster than the \emph{observed} rate at which new cases are generated, which does not depend on the amount of early transmission.
On the other hand, the original explanation of the result relied on a strength-based arugment: increasing the amount of early transmission decreases the basic reproductive number, which negatively correlates with the outcome of the ART intervention \citep{eaton2014proportion}.
While both strength- and speed-based frameworks give the same conclusion about the outcome of an intervention, one can provide a clearer interpretation of the outcome than the other, depending on the situation.
